---
title: "Figure 1 and associated groupings"
author: "Sven Tobias-Hunefeldt"
date: "30/03/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Set up
Import data and refine as done using the set up script
```{r Install ecotone assigner}

install.packages("/YOUR/PATH/HERE/EcotoneFinder_0.2.0.tar.gz", repos = NULL, type = "source")

```
```{r load packages}

library(EcotoneFinder)
library(colorspace)
library(Rmisc)
library(reshape2)
library(phyloseq)
library(ggpubr)
library(ggplot2)
library(cluster)

```
#Set up dataframes for ecotone divider
##Prokaryotes
```{r Prokaryotes}

#Restore a pristine copy from back up
Phyloseq_Biofilm_16S = Phyloseq_Biofilm_16S_v0

#Look otu table and metadata, are they mergable?
rownames(Phyloseq_Biofilm_16S@sam_data)
rownames(Phyloseq_Biofilm_16S@otu_table)
#Merge them
OTU_Meta_df_16S = merge(Phyloseq_Biofilm_16S@sam_data, Phyloseq_Biofilm_16S@otu_table, by = 0)
#Reorder factors for downstream analysis
OTU_Meta_df_16S$Sample_time = factor(OTU_Meta_df_16S$Sample_time, 
                                    levels = c("0", "7", "14", "19", "28", "42", "56"))

#Subset merged dataframe
JustBiofilm_test = subset(OTU_Meta_df_16S, Substrate == "Acryl" | Substrate == "Glass" | Substrate == "Tile" | Substrate == "Pine")

#Assign informative rownames to dataframe
rownames(JustBiofilm_test) = JustBiofilm_test$Row.names
#Check dataframe to see what needs removing
JustBiofilm_test[c(1:5), c(1:20)]
#Remove excess information
JustBiofilm_test[,c(1:12,14,15)] = NULL #Only keep sample time
#Check dataframe to see if it worked
JustBiofilm_test[c(1:5), c(1:20)]
#Melt into long format
JustBiofilm_long = melt(JustBiofilm_test, id.vars = "Sample_time")
#Check for errors
JustBiofilm_long[c(1:239), c(1:3)]

#Identify type of variable for downstream anlysis
typeof(JustBiofilm_long$Sample_time)
typeof(JustBiofilm_long$variable)
typeof(JustBiofilm_long$value)
#Change data type for easier analysis
JustBiofilm_long$Sample_time = as.numeric(as.character(JustBiofilm_long$Sample_time))
JustBiofilm_long$variable = as.factor(as.character(JustBiofilm_long$variable))
JustBiofilm_long$value = as.numeric(as.character(JustBiofilm_long$value))
#Check to see if it worked
typeof(JustBiofilm_long$Sample_time)
typeof(JustBiofilm_long$variable)
typeof(JustBiofilm_long$value)

#Make summary for identify trends
JustBiofilm_summary = summarySE(JustBiofilm_long, measurevar = "value", groupvars = c("Sample_time", "variable"))
#Check for errors
JustBiofilm_summary[c(1:100), c(1:7)]
test = "ASV3"
filter(JustBiofilm_summary, JustBiofilm_summary$variable %in% test)

#Convert data back into a wide format
JustBiofilm_wide_16S = reshape2::dcast(JustBiofilm_summary, Sample_time ~ variable, value.var = "value")
#Check for error
JustBiofilm_wide_16S[c(1:7), c(1:100)]

#Change names for downstream analysis
colnames(JustBiofilm_wide_16S) = gsub("ASV", "Sp.", colnames(JustBiofilm_wide_16S))
colnames(JustBiofilm_wide_16S) = gsub("Sample_time", "Distance", colnames(JustBiofilm_wide_16S))
#Check for errors
JustBiofilm_wide_16S[c(1:7), c(1:100)]
#Redefine for easier analysis
JustBiofilm_wide_16S$Distance <- factor(JustBiofilm_wide_16S$Distance,
                         levels=(unique(JustBiofilm_wide_16S$Distance)))

```
##Eukaryotes
```{r Eukaryotes}

#Restore a pristine copy from back up
Phyloseq_Biofilm_18S = Phyloseq_Biofilm_18S_v0

#Look otu table and metadata, are they mergable?
rownames(Phyloseq_Biofilm_18S@sam_data)
rownames(Phyloseq_Biofilm_18S@otu_table)
#Merge them
OTU_Meta_df_18S = merge(Phyloseq_Biofilm_18S@sam_data, Phyloseq_Biofilm_18S@otu_table, by = 0)
#Reorder factors for downstream analysis
OTU_Meta_df_18S$Sample_time = factor(OTU_Meta_df_18S$Sample_time, 
                                    levels = c("0", "7", "14", "19", "28", "42", "56"))

#Subset merged dataframe
JustBiofilm_test = subset(OTU_Meta_df_18S, Substrate == "Acryl" | Substrate == "Glass" | Substrate == "Tile" | Substrate == "Pine")

#Assign informative rownames to dataframe
rownames(JustBiofilm_test) = JustBiofilm_test$Row.names
#Check dataframe to see what needs removing
JustBiofilm_test[c(1:5), c(1:20)]
#Remove excess information
JustBiofilm_test[,c(1:12,14,15)] = NULL #Only keep sample time
#Check dataframe to see if it worked
JustBiofilm_test[c(1:5), c(1:20)]
#Melt into long format
JustBiofilm_long = melt(JustBiofilm_test, id.vars = "Sample_time")
#Check for errors
JustBiofilm_long[c(1:239), c(1:3)]

#Identify type of variable for downstream anlysis
typeof(JustBiofilm_long$Sample_time)
typeof(JustBiofilm_long$variable)
typeof(JustBiofilm_long$value)
#Change data type for easier analysis
JustBiofilm_long$Sample_time = as.numeric(as.character(JustBiofilm_long$Sample_time))
JustBiofilm_long$variable = as.factor(as.character(JustBiofilm_long$variable))
JustBiofilm_long$value = as.numeric(as.character(JustBiofilm_long$value))
#Check to see if it worked
typeof(JustBiofilm_long$Sample_time)
typeof(JustBiofilm_long$variable)
typeof(JustBiofilm_long$value)

#Make summary for identify trends
JustBiofilm_summary = summarySE(JustBiofilm_long, measurevar = "value", groupvars = c("Sample_time", "variable"))
#Check for errors
JustBiofilm_summary[c(1:100), c(1:7)]
test = "ASV3"
filter(JustBiofilm_summary, JustBiofilm_summary$variable %in% test)

#Convert data back into a wide format
JustBiofilm_wide_18S = reshape2::dcast(JustBiofilm_summary, Sample_time ~ variable, value.var = "value")
#Check for error
JustBiofilm_wide_18S[c(1:7), c(1:100)]

#Change names for downstream analysis
colnames(JustBiofilm_wide_18S) = gsub("ASV", "Sp.", colnames(JustBiofilm_wide_18S))
colnames(JustBiofilm_wide_18S) = gsub("Sample_time", "Distance", colnames(JustBiofilm_wide_18S))
#Check for errors
JustBiofilm_wide_18S[c(1:7), c(1:100)]
#Redefine for easier analysis
JustBiofilm_wide_18S$Distance <- factor(JustBiofilm_wide_18S$Distance,
                         levels=(unique(JustBiofilm_wide_18S$Distance)))

```

#Run analyses for ecotone divider
```{r Make palette}
mock.df = data.frame(Dissimilarity = c(0,
                                       100),
                     Sample = c(1,
                                2),
                     ygroup = c(100,
                                10))
heatmap(mock.df, aes(x = as.numeric(Sample), y = ygroup, fill = Dissimilarity))+
  geom_tile()+
  scale_color_continuous(values())

```

##Prokaryotes
```{r}
library("RColorBrewer")
col <- colorRampPalette(brewer.pal(8, "Dark2"))(256)

#Identify best number of clusters
DistEco(data = type.convert(JustBiofilm_wide_16S[,-1]),
        distance.method = "sorensen",
        transpose = F, #Calculate groups as sites
        spinglass = T, #Use spinglass algorithm to produce statistical groups for the network
        plot = "heatmap",
        run = 10, #How often to repeat the spinglass algorithm
        spinglass.groups = c("raw")) #The type of grouping to use from the results of the spinglass algorithm


tiff("/YOUR/PATH/HERE/Figures_Tables/ProkHeat.tiff", height = 6.1, width = 8, units = "in", res = 300)
DistEco(data = type.convert(JustBiofilm_wide_16S[,-1]),
        distance.method = "sorensen",
        transpose = F, #Calculate groups as sites
        spinglass = T, #Use spinglass algorithm to produce statistical groups for the network
        plot = "heatmap",
        run = 10, #How often to repeat the spinglass algorithm
        spinglass.groups = c("raw")) #The type of grouping to use from the results of the spinglass algorithm
dev.off()

DistEco(data = type.convert(JustBiofilm_wide_16S[,-1]),
        distance.method = "sorensen",
        transpose = F, #Calculate groups as sites
        spinglass = T, #Use spinglass algorithm to produce statistical groups for the network
        plot = "network",
        run = 1000, #How often to repeat the spinglass algorithm
        spinglass.groups = c("raw")) #The type of grouping to use from the results of the spinglass algorithm

#### 2 cluster Analyses####
SyntheticEcoFinder_2_16S <- EcotoneFinder(data = type.convert(JustBiofilm_wide_16S[,-1]),
                                     dist = type.convert(JustBiofilm_wide_16S$Distance),
                                     method = "vegclust",
                                     groups = 2, 
                                     standardize = "hellinger",
                                     diversity = "all")
#Extract useful info for plotting
Forplot_2_16S = SyntheticEcoFinder_2_16S$vegclust$memb
#Generate plot
Plot_2_16S_eco = ggplot(Forplot_2_16S, 
                    aes(x = as.factor(c(7, 14, 19, 28, 42, 56))))+
  geom_line(aes(y = M1, 
                colour = "red", 
                group = "M1"), size = 2)+
  geom_line(aes( y = M2, 
                 colour = "magenta", 
                 group = "M2"), size = 2)+
  xlab("Time (days)")+
  ylab("Ecological gradient based \non Fuzzy Cluster Means")+
  theme_bw()+
  My_Theme+
  theme(axis.text.y = element_blank(),
        legend.position = "none")
#Check plot
Plot_2_16S_eco


#### 3 cluster Analyses####
SyntheticEcoFinder_3_16S <- EcotoneFinder(data = type.convert(JustBiofilm_wide_16S[,-1]),
                                     dist = type.convert(JustBiofilm_wide_16S$Distance),
                                     method = "vegclust",
                                     groups = 3, 
                                     standardize = "hellinger",
                                     diversity = "all")
#Extract useful info for plotting
Forplot_3_16S = SyntheticEcoFinder_3_16S$vegclust$memb
#Generate plot
Plot_3_16S = ggplot(Forplot_3_16S, 
                    aes(x = as.factor(c(7, 14, 19, 28, 42, 56))))+
  geom_line(aes(y = M1, 
                colour = "red", 
                group = "M1"), size = 2)+
  geom_line(aes( y = M2, 
                 colour = "green", 
                 group = "M2"), size = 2)+
  geom_line(aes( y = M3, 
                 colour = "magenta", 
                 group = "M3"), size = 2)+
  xlab("Time (days)")+
  ylab("Ecological gradient based \non Fuzzy Cluster Means")+
  theme_bw()+
  My_Theme+
  theme(axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")
#Check plot
Plot_3_16S


#### 4 cluster Analyses####
SyntheticEcoFinder_4_16S <- EcotoneFinder(data = type.convert(JustBiofilm_wide_16S[,-1]),
                                     dist = type.convert(JustBiofilm_wide_16S$Distance),
                                     method = "vegclust",
                                     groups = 4, 
                                     standardize = "hellinger",
                                     diversity = "all")
#Extract useful info for plotting
Forplot_4_16S = SyntheticEcoFinder_4_16S$vegclust$memb
#Generate plot
Plot_4_16S = ggplot(Forplot_4_16S, 
                    aes(x = as.factor(c(7, 14, 19, 28, 42, 56))))+
  geom_line(aes(y = M1, 
                colour = "red", 
                group = "M1"), size = 2)+
  geom_line(aes( y = M2, 
                 colour = "green", 
                 group = "M2"), size = 2)+
  geom_line(aes( y = M3, 
                 colour = "magenta", 
                 group = "M3"), size = 2)+
    geom_line(aes( y = M4, 
                 colour = "blue", 
                 group = "M4"), size = 2)+
  xlab("Time (days)")+
  ylab("Ecological gradient based \non Fuzzy Cluster Means")+
  theme_bw()+
  My_Theme+
  theme(axis.text.y = element_blank(),
        legend.position = "none")
#Check plot
Plot_4_16S

```
##Eukaryotes
```{r}
#Identify best number of clusters
DistEco(data = type.convert(JustBiofilm_wide_18S[,-1]),
        distance.method = "sorensen",
        transpose = F, #Calculate groups as sites
        spinglass = T, #Use spinglass algorithm to produce statistical groups for the network
        plot = "heatmap",
        run = 10, #How often to repeat the spinglass algorithm
        spinglass.groups = c("raw")) #The type of grouping to use from the results of the spinglass algorithm

tiff("/Volumes/Lab_book/Biofilm_project/Analysis/Clean_ms_code/Reanalysis/Figures_Tables/EukHeat.tiff", height = 6.1, width = 8, units = "in", res = 300)
DistEco(data = type.convert(JustBiofilm_wide_18S[,-1]),
        distance.method = "sorensen",
        transpose = F, #Calculate groups as sites
        spinglass = T, #Use spinglass algorithm to produce statistical groups for the network
        plot = "heatmap",
        run = 10, #How often to repeat the spinglass algorithm
        spinglass.groups = c("raw")) #The type of grouping to use from the results of the spinglass algorithm
dev.off()

DistEco(data = type.convert(JustBiofilm_wide_18S[,-1]),
        distance.method = "sorensen",
        transpose = F, #Calculate groups as sites
        spinglass = T, #Use spinglass algorithm to produce statistical groups for the network
        plot = "network",
        run = 100, #How often to repeat the spinglass algorithm
        spinglass.groups = c("raw")) #The type of grouping to use from the results of the spinglass algorithm

#### 2 cluster Analyses####
SyntheticEcoFinder_2_18S <- EcotoneFinder(data = type.convert(JustBiofilm_wide_18S[,-1]),
                                     dist = type.convert(JustBiofilm_wide_18S$Distance),
                                     method = "vegclust",
                                     groups = 2, 
                                     standardize = "hellinger",
                                     diversity = "all")
#Extract useful info for plotting
Forplot_2_18S = SyntheticEcoFinder_2_18S$vegclust$memb
#Generate plot
Plot_2_18S_eco = ggplot(Forplot_2_18S, 
                    aes(x = as.factor(c(7, 14, 19, 28, 42, 56))))+
  geom_line(aes(y = M1, 
                colour = "red", 
                group = "M1"), size = 2)+
  geom_line(aes( y = M2, 
                 colour = "magenta", 
                 group = "M2"), size = 2)+
  xlab("Time (days)")+
  ylab("Ecological gradient based on Fuzzy Cluster Means")+
  theme_bw()+
  My_Theme+
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none")
#Check plot
Plot_2_18S_eco


#### 3 cluster Analyses####
SyntheticEcoFinder_3_18S <- EcotoneFinder(data = type.convert(JustBiofilm_wide_18S[,-1]),
                                     dist = type.convert(JustBiofilm_wide_18S$Distance),
                                     method = "vegclust",
                                     groups = 3, 
                                     standardize = "hellinger",
                                     diversity = "all")
#Extract useful info for plotting
Forplot_3_18S = SyntheticEcoFinder_3_18S$vegclust$memb
#Generate plot
Plot_3_18S = ggplot(Forplot_3_18S, 
                    aes(x = as.factor(c(7, 14, 19, 28, 42, 56))))+
  geom_line(aes(y = M1, 
                colour = "red", 
                group = "M1"), size = 2)+
  geom_line(aes( y = M2, 
                 colour = "green", 
                 group = "M2"), size = 2)+
  geom_line(aes( y = M3, 
                 colour = "magenta", 
                 group = "M3"), size = 2)+
  xlab("Time (days)")+
  ylab("Ecological gradient based on Fuzzy Cluster Means")+
  theme_bw()+
  My_Theme+
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")
#Check plot
Plot_3_18S


#### 4 cluster Analyses####
SyntheticEcoFinder_4_18S <- EcotoneFinder(data = type.convert(JustBiofilm_wide_18S[,-1]),
                                     dist = type.convert(JustBiofilm_wide_18S$Distance),
                                     method = "vegclust",
                                     groups = 4, 
                                     standardize = "hellinger",
                                     diversity = "all")
#Extract useful info for plotting
Forplot_4_18S = as.data.frame(SyntheticEcoFinder_4_18S$vegclust$memb)
#Generate plot
Plot_4_18S = ggplot(Forplot_4_18S, 
                    aes(x = as.factor(c(7, 14, 19, 28, 42, 56))))+
  geom_line(aes(y = M1, 
                colour = "red", 
                group = "M1"), size = 2)+
  geom_line(aes( y = M2, 
                 colour = "green", 
                 group = "M2"), size = 2)+
  geom_line(aes( y = M3, 
                 colour = "magenta", 
                 group = "M3"), size = 2)+
    geom_line(aes( y = M4, 
                 colour = "blue", 
                 group = "M4"), size = 2)+
  xlab("Time (days)")+
  ylab("Ecological gradient based on Fuzzy Cluster Means")+
  theme_bw()+
  My_Theme+
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none")
#Check plot
Plot_4_18S

```
#Prokaryotic silhouette analysis
Two clusters shows the lowest number of samples below 0 (9 vs 3's 10) while also retaining a high intra-group value
```{r}

#Restore a pristine copy from back up
Phyloseq_Biofilm_16S = Phyloseq_Biofilm_16S_v0

#Subset to only substrate associated biofilms
Phyloseq_Biofilm_16S_noA = subset_samples(Phyloseq_Biofilm_16S, Project == "Marine_Biofilm" & Sample_type == "Biofilm" & !Substrate == "Water"& !Substrate == "Mesh")

#Extract otu table
otutable_df_16S = as.data.frame(as.matrix(otu_table(Phyloseq_Biofilm_16S_noA)))
nrow(otutable_df_16S)

#2 clusters
test2_16S = pam(otutable_df_16S, k = 2)
Plot2_16S = recordPlot(plot(silhouette(test2_16S)))
#This shows minimal number of misclassified samples

#Save figure
tiff("/YOUR/PATH/HERE/Figures_Tables/SupplementaryFigure1msC_ProkaryoticSilhouette.tiff", height = 6.66, width = 8, units = "in", res = 300)
test2_16S = pam(otutable_df_16S, k = 2)
Plot2_16S = recordPlot(plot(silhouette(test2_16S)))
dev.off()

test2_16S$clustering[test2_16S$clustering == 1]
31/111 #Proportion that is from day 7 and 14

test2_16S$clustering[test2_16S$clustering == 2]
9/12 #Proportion that is from day 7 and 14

#3 clusters
test3_16S = pam(otutable_df_16S, k = 3)
Plot3_16S = recordPlot(plot(silhouette(test3_16S)))

test3_16S$clustering[test3_16S$clustering == 1]
25/100 #Rest is made up of other mix of samples

test3_16S$clustering[test3_16S$clustering == 2]
9/12 #Mix of lots of day 19 samples

test3_16S$clustering[test3_16S$clustering == 3]
6/11 #Rest is made up of day 28-42 samples

#4 clusters
test4_16S = pam(otutable_df_16S, k = 4)
Plot4_16S = recordPlot(plot(silhouette(test4_16S)))

test4_16S$clustering[test4_16S$clustering == 1]
11/14 #Proportion that is from day 7 and 14

test4_16S$clustering[test4_16S$clustering == 2]
14/35 #Proportion that is from day 7 and 14

test4_16S$clustering[test4_16S$clustering == 3]
9/63 #Proportion that is from day 7 and 14

test4_16S$clustering[test4_16S$clustering == 4]
6/11 #Proportion that is from day 7 and 14

#5 clusters
test5_16S = pam(otutable_df_16S, k = 5)
Plot5_16S = recordPlot(plot(silhouette(test5_16S)))

#6 clusters
test6_16S = pam(otutable_df_16S, k = 6)
Plot6_16S = recordPlot(plot(silhouette(test6_16S)))

```
#Eukaryotic silhouette analysis
No seeming cluster effect - assessing day 7 clusteirng from everything else based on other evidence
```{r}

#Restore a pristine copy from back up
Phyloseq_Biofilm_18S = Phyloseq_Biofilm_18S_v0

#Subset to only substrate associated biofilms
Phyloseq_Biofilm_18S_noA = subset_samples(Phyloseq_Biofilm_18S, Project == "Marine_Biofilm" & Sample_type == "Biofilm" & !Substrate == "Water"& !Substrate == "Mesh")

#Extract otu table
otutable_df_18S = as.data.frame(as.matrix(otu_table(Phyloseq_Biofilm_18S_noA)))
nrow(otutable_df_18S)

#2 clusters
test2_18S = pam(otutable_df_18S, k = 2)
Plot2_18S = recordPlot(plot(silhouette(test2_18S)))
#This shows minimal number of misclassified samples

#Save figure
tiff("/YOUR/PATH/HERE/Figures_Tables/SupplementaryFigure1msD_EukaryoticSilhouette.tiff", height = 6.66, width = 8, units = "in", res = 300)
Plot2_18S = recordPlot(plot(silhouette(test2_18S)))
dev.off()

test2_18S$clustering[test2_18S$clustering == 1]
38/110 #Proportion that is from day 7 and 14

test2_18S$clustering[test2_18S$clustering == 2]
3/20 #Proportion that is from day 7 and 14

#3 clusters
test3_18S = pam(otutable_df_18S, k = 3)
Plot3_18S = recordPlot(plot(silhouette(test3_18S)))

test3_18S$clustering[test3_18S$clustering == 1]
7/7 #Pure day 7 samples

test3_18S$clustering[test3_18S$clustering == 2]
12/103 #Mix of lots of day 14-56 samples

test3_18S$clustering[test3_18S$clustering == 3]
0/20 #Rest is made up of day 28-42 samples

#4 clusters
test4_18S = pam(otutable_df_18S, k = 4)
Plot4_18S = recordPlot(plot(silhouette(test4_18S)))

test4_18S$clustering[test4_18S$clustering == 1]
7/7 #Proportion that is from day 7

test4_18S$clustering[test4_18S$clustering == 2]
12/95 #Proportion that is from day 7

test4_18S$clustering[test4_18S$clustering == 3]
0/15 #Proportion that is from day 7 

test4_18S$clustering[test4_18S$clustering == 4]
0/13 #Proportion that is from day 7 and 14

#5 clusters
test5_18S = pam(otutable_df_18S, k = 5)
Plot5_18S = recordPlot(plot(silhouette(test5_18S)))

#6 clusters
test6_18S = pam(otutable_df_18S, k = 6)
Plot6_18S = recordPlot(plot(silhouette(test6_18S)))

```

#Make figure using ggarrange

```{r}
ggEcotone(Prok_groupings_heat)

#Make figure
SuppFigure3 = ggarrange(Plot_2_16S_eco, Plot_2_18S_eco,
                        F, F,
                        common.legend = F,
          ncol = 2,
          nrow = 2)
#Check for error
SuppFigure3

#Save figure
tiff("/YOUR/PATH/HERE/Figures_Tables/SupplementaryFigure3ms_EcotoneAndSilhouette.tiff", height = 12.2, width = 16, units = "in", res = 300)
SuppFigure3
dev.off()

```

